### 2017-05-11 papers/paper.tex
Adds logs, improves docs, extends paper

\documentclass{article}

\usepackage{indentfirst}
\usepackage{setspace}
\doublespacing

% ================================================================================= 
% Package for showing source code
% =================================================================================

\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
language=C,
aboveskip=.5mm,
belowskip=.5mm,
showstringspaces=false,
columns=flexible,
basicstyle={\scriptsize\ttfamily},
numbers=none,
numberstyle=\tiny\color{gray},
keywordstyle=\color{blue},
commentstyle=\color{dkgreen},
stringstyle=\color{mauve},
breaklines=false,
breakatwhitespace=true,
tabsize=3
}

% ================================================================================= 
% Package for flowcharts/diagrams
% =================================================================================

\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{startstop} = [rectangle, rounded corners, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=red!30]
\tikzstyle{io}        = [rectangle, minimum width=3cm, minimum height=1cm,text centered, draw=black, fill=blue!30]
\tikzstyle{process}   = [diamond, minimum width=2cm, minimum height=0cm, text centered, draw=black, fill=orange!30]
\tikzstyle{arrow}     = [thick,->,>=stealth]

% ==================================================
% Paper
% ==================================================

\title{Quantum Programming Interface for NP-complete problems}
\date{05-10-2017}
\author{Lucas Saldyt}

\begin{document}

\maketitle
\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}

% ==================================================
\section{Abstract}
% ==================================================

Programming a quantum annealer to solve certain kinds of NP-complete problems is actually very simple.
Because of the rich software toolchains surrounding the DWave, it is easy to describe problems for the annealer to solve.
This paper compares various implementations of the Knapsack problem: Classically, using the DWave, and using the IBM Quantum Experience.
Formal benchmarking is not a direct part of report. Formal benchmarks have been performed in [citations], and will be taken into account in the analysis of this paper.
[Conclusion recap]

% ==================================================
\section{Introduction}
% ==================================================

The Knapsack problem and other NP-complete problems have immediate real-world applications. 
However, as the size of an NP-complete problem increases, it becomes very difficult to solve.
Quantum Computation can allow NP-complete problems to be solved with a polynomial speedup.
This paper explores different interfaces in the hope of improving the availability of quantum computation, and demonstrating the ease of use of existing interfaces.

\newpage

% ==================================================
\section{Methods}
% ==================================================

% --------------------------------------------------
\subsection{Classical Knapsack}
% --------------------------------------------------

Traditionally, the optimal way to solve the knapsack is through dynamic programming.
However, the code necessary is potentially difficult to write from scratch:

\lstset{language=Python}
\begin{lstlisting}
import functools

def memoize(obj):
    '''
    Decorator that caches function calls, allowing dynamic programming.
    If a function is called twice with the same arguments, the cached result is used.

    obj: function to be cached
    '''
    cache = obj.cache = {}
    @functools.wraps(obj)
    def memoizer(*args, **kwargs):
        key = str(args) + str(kwargs)
        if key not in cache:
            cache[key] = obj(*args, **kwargs)
        return cache[key]
    return memoizer

def knapsack(items, outerConstraints):
    '''
    Solve the knapsack problem

    `items`: a sequence of pairs `(value, weight, volume)`, where `value` is
      a number and `weight` is a non-negative integer.

    `outerConstraints`: a list of numbers representing the maximum values 
      for each respective constraint

    `return`: a pair whose first element is the sum of values in the most
    valuable subsequence, and whose second element is the subsequence.
    '''

    # Return the value of the most valuable subsequence of the first i
    # elements in items whose constraints are satisfied
    @memoize
    def bestvalue(i, constraints): # j is a constraint
        if i == 0: return 0
        value, *limiters = items[i - 1]
        tests = [l > v for l, v in zip(limiters, constraints)]
        if any(tests): # constraint checking
            return bestvalue(i - 1, constraints)
        else:          # maximizing
            modifications = [v - l for l, v in zip(limiters, constraints)]
            return max(bestvalue(i - 1, constraints),
                       bestvalue(i - 1, modifications) + value)

    # Create and modify a copy of
    k = outerConstraints
    result = []
    for i in range(len(items), 0, -1):
        if bestvalue(i, k) != bestvalue(i - 1, k):
            result.append(items[i - 1])
            k = [c - items[i - 1][n + 1] for n, c in enumerate(k)]
    result.reverse()

    print('Knapsack problem solved classically with {} static evaluations:'.format(len(bestvalue.cache)))
    choices = [items.index(choice) for choice in result]
\end{lstlisting}

\newpage

A simpler, naive implementation of the knapsack problem would try every combination of items, choosing the one with the best value that satisfied the constraints.

\lstset{language=Python}
\begin{lstlisting}
def naive(items, capacities):
    selection = []
    bestval   = 0
    for comb in powerset(items):
        if len(comb) > 0:
            setcaps = setvals(comb)
            for j, capacity in enumerate(capacities):
                if j == 0:
                    if bestval > setcaps[0]:
                        break
                else:
                    if setcaps[j] > capacity:
                        break
                if j == len(capacities) - 1: # All constraints satisfied for this set
                    selection = comb
    choices = [items.index(choice) for choice in selection]
    return bestval, selection, choices
\end{lstlisting}

\newpage

However, a fully polynomial approximation scheme is more accurately comparable to the problem being run on the DWave.
A FPTAS simplifies the problem given, solving it to withing a margin of error.
After simplification, FPTAS uses dynamic programming.
For problems with few variables, the results of FPTAS are very similar to the results of dynamic programming, even if the margin of error is allowed to be high.

\lstset{language=Python}
\begin{lstlisting}
def fptas(items, capacities, e=0.1):
    maxvalue = max(items, key=itemgetter(0))[0]
    k = e * (maxvalue / len(items))
    items = [(v/k, *rest) for v, *rest in items]
    return knapsack(items, capacities)
\end{lstlisting}

\newpage

Below is output from all three classical solvers on an eight variable knapsack problem:

\begin{verbatim}
fptas : 0.0007245789747685194
{'G', 'B', 'D', 'H', 'E'}
value is satisfied (70)
weight is satisfied (60)
volume is satisfied (55)
________________________________________________________________________________

naive : 0.0013742649462074041
{'G', 'B', 'D', 'H', 'E'}
value is satisfied (70)
weight is satisfied (60)
volume is satisfied (55)
________________________________________________________________________________

dynamic: 0.0006771821063011885
{'G', 'B', 'D', 'H', 'E'}
value is satisfied (70)
weight is satisfied (60)
volume is satisfied (55)
________________________________________________________________________________
\end{verbatim}

Above 8 variables, the naive implementation times out or returns an incorrect answer.

\newpage

QMASM solving: Verilog and Prolog

\lstset{language=Verilog}
\begin{lstlisting}
test
\end{lstlisting}

\lstset{language=Prolog}
\begin{lstlisting}
test
\end{lstlisting}

\newpage

IBM Quantum experience

% ==================================================
\section{Discussion}
% ==================================================

Which interface is simplest? Easiest to use?
Simple interface must be justified: If it isn't useful for solving problems, then the simplicity of the interface doesn't matter.
Even if a simpler interface works for certain kinds of problems, be careful when making claims.

Cite other benchmarks, don't do many.

% ==================================================
\section{Appendix}
% ==================================================

\begin{thebibliography}{9}
    \bibitem{gps}
        A. Newell, J. C. Shaw, H. A. Simon
        \textit{Report on a general problem-solving program}
        The Rand corporation, Santa Monica, California, 1958
\end{thebibliography}

\end{document}