### 2017-05-12 papers/essay.tex
Adds essay outline

\documentclass{article}

\usepackage{indentfirst}
\usepackage{setspace}
\doublespacing

% ==================================================
% Paper
% ==================================================

\title{Senior Experience Reflective Essay}
\date{05-11-2017}
\author{Lucas Saldyt}

\begin{document}

\maketitle
\pagenumbering{gobble}
\newpage
\pagenumbering{arabic}

% ==================================================
\section{Reflections}
% ==================================================

• What else have you learned about yourself? About your project topic?
• Is the project topic something you are likely to pursue once senior projects are over?
• Without a quiz, paper, test, homework assignment or teacher to motivate you, what has
motivated you to work on your senior project this week? Is it purely getting through the
last graduation requirement, or do you truly enjoy spending time on this project that you
created? Is it hard to motivate yourself to spend time on the project? Write some ideas
about what motivates you.
• If your project were a class and you were the teacher, how would you assess/test/evaluate
your students’ progress, understanding, skill? Would you give a written test based on
knowledge, a hands-on demonstration of skill, or some other evaluation?
• If you made a soundtrack to your senior project, what songs would you include and why?
• Describe a metaphor for this experience. Fill in the blank and elaborate: This experience
has been like...
• What advice would you give next year’s seniors?

% ==================================================
\section{Appendix}
% ==================================================

\end{document}